Busch, Khakhalin. Intrinsic temporal tuning

### 2018-12-08 reviews and responses.md

> 6.	Figure 2B Amplitude is misspelled.

Fixed.

> 7.	The narrative logic on page 9 lines 11-22 is confusing. How does homeostatic plasticity, if it truly is at play here, "increase their ability to respond to unusual patterns of synaptic activation." If anything wouldn't homeostasis decrease dynamic range by pushing all cells toward a central set point?

We agree that the word "homeostatic" was confusing in this context. We now simplified the sentence:

@@

> The paper is, in my opinion, hard to read because the authors are often not clear about what they mean, refer to previous parts of the paper assuming the reader has a perfect memory or is willing to turn back and forth across pages, and use non-traditional terminology. In addition, the statistically analysis appears to be pooling many data points from non-independent cells, rendering many of the conclusions invalid. Further, the experiments intended to measure intrinsic properties (in Figures 1 and 2) appear to have been performed without synaptic blockers, so the the adaptation that they measure could be in the circuit (inhibitory recruitment) instead of in the intrinsic properties.

> There is indeed some merit in the goals of the study (to see if adaptation changes with experience in addition to conductance/response gains), but it has so many technical problems at present that it would need to be radically different (perhaps with experiments and analysis re-done) to perhaps be of interest to the broad readership of J Neuroscience.

&nbsp;

On references to previous parts, thank you for pointing it out! We now tried to link different parts of the paper more tightly together, by reminding the reader about previously made statements every time we make a reference.

On statistical analysis: we do not pool data points from non-independent cells, but used a repeated measures analysis, verified by mixed model analysis (see Methods, as well as corrections described in our responses to Reviewer 1).

On synaptic blockers: it is true that we measured intrinsic properties without synaptic blockers, but this is considered a standard practice in this preparation, as most earlier studies measured IV curves without synaptic blockers (Aizenman .. Cline 2002; Aizenman .. Cline 2003; James .. Aizenman 2015; Ciarleglio Khakhalin .. Aizenman 2015). What makes this practice possible is a relatively weak recruitment of polysynaptic activity after activation of a single tectal cell (Bell .. Aizenman 2011).

### 2018-12-08 reviews and responses.md

F(1,677)=30.4, p=5e-8, n cells=28, 29. Here and below, when F-values are reported, we use 4-way fixed effects analysis of variance with selected interactions, and cell id included as repeated measures factor, see "Methods" for a detailed description.

> 5.	Page 5 line 19 "This suggests that unlike the change in overall intrinsic excitability, the average temporal retuning was "adaptive" rather than "homeostatic", as neurons exposed to shorter stimuli (flashes) became more equipped to process shorter activation patterns." While it is appealing to give a name to this form of plasticity, the looming stimulus, which can be thought of as slower than the flash nonetheless caused a similar type of shift in response preference in the neurons, which runs counter the notion that there exists a meaningful dichotomy of adaptive vs. homeostatic change here.

We think that there exists a certain dichotomy here, but we agree that our description of it was not clear enough. We now replaced it with the following, longer exposition:

"... This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, follow two different kinds of logic. The overall excitability is homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as “adaptive”, as neurons exposed to shorter stimuli (flashes) became relatively *more* responsive to shorter stimuli, and less responsive to longer stimuli, which is opposite of what one would expect for a homeostatic retuning. We chose to call this type of plasticity "adaptive", as presumably it means that after exposure to faster stimuli, neurons become more equipped to process faster patterns of activation."

### 2018-12-08 main.tex
minor: homeostatic passage

Across input conductances of different lengths (100, 200, 500, and 1000 ms), neurons exposed to flashes and looming stimuli responded differently than in control (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs typically evoked stronger spiking, whereas neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike number for longer conductance injections (Fig 1E). In essence, while control neurons “preferred” longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, each follow different logic, and cannot be easily described by a blanket statement. The overall excitability is "homeostatic", as in response to stronger stimulation neurons became less excitable. The temporal retuning however can be better described as “adaptive”, as neurons exposed to shorter stimuli (flashes) became relatively more responsive to shorter stimuli, and so presumably more equipped to process faster activation patterns.

@@

To sum up, we present a novel case of temporal selectivity in non-oscillatory neurons in a distributed sensory network, and demonstrate that intrinsic temporal tuning of neural cells correlates with their synaptic properties. It is particularly interesting that the temporal co-tuning we observed was homeostatic in nature, as cells tended to be selective for inputs of dynamics they did not usually experience. We hypothesize that this homeostatic adjustment of temporal tuning is the reason why this tuning was so easily disrupted by strong visual stimulation, when for a few hours we drastically changed the statistics of inputs received by every cell. Tuning to "unusual stimuli" at the level of individual neurons fits into the narrative of information transfer maximization \citep{brenner2000} and network criticality \citep{rubinov2011}, wherein every element of a network tries to locally maximize its influence over the overall computation. This may have intriguing consequences for the function and development of sensory networks in the brain, which can be further probed by computational modeling \citep{khakhalin2014,jang2016}, and verified in experiments. We also suspect that any cell with a sufficiently large dendritic tree would be able to tune its temporal intrinsic selectivity the way we describe. We expect that in most cases these changes would not be noticeable in experiments with standard voltage- and current-clamp protocols, but can be probed with a dynamic clamp technique. It would therefore be very interesting to see whether our results will replicate in other sensory systems, such as the mammalian cortex.

>>>>

Across input conductances of different lengths (100, 200, 500, and 1000 ms), neurons exposed to flashes and looming stimuli responded differently than in control (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs typically evoked stronger spiking, whereas neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike number for longer conductance injections (Fig 1E). In essence, while control neurons “preferred” longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, follow two different kinds of logic. The overall excitability is homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as “adaptive”, as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, which is opposite of what one would expect for a fully homeostatic retuning. We chose to call this type of plasticity "adaptive", as presumably it means that after exposure to faster stimuli, neurons become more equipped to process faster patterns of activation.

@@

To sum up, we present a novel case of temporal selectivity in non-oscillatory neurons in a distributed sensory network, and demonstrate that intrinsic temporal tuning of neural cells correlates with their synaptic properties. It is particularly interesting that the temporal co-tuning we observed was homeostatic in nature, as cells tended to be selective for inputs of dynamics they did not usually experience. We hypothesize that this adjustment of temporal tuning is the reason why this tuning was so easily disrupted by strong visual stimulation, when for a few hours we drastically changed the statistics of inputs received by every cell. Tuning to "unusual stimuli" at the level of individual neurons fits into the narrative of information transfer maximization \citep{brenner2000} and network criticality \citep{rubinov2011}, wherein every element of a network tries to locally maximize its influence over the overall computation. This may have intriguing consequences for the function and development of sensory networks in the brain, which can be further probed by computational modeling \citep{khakhalin2014,jang2016}, and verified in experiments. We also suspect that any cell with a sufficiently large dendritic tree would be able to tune its temporal intrinsic selectivity the way we describe. We expect that in most cases these changes would not be noticeable in experiments with standard voltage- and current-clamp protocols, but can be probed with a dynamic clamp technique. It would therefore be very interesting to see whether our results will replicate in other sensory systems, such as the mammalian cortex.
